Model Adaptivity for Goal-Oriented Inference using Adjoints

### 2017-06-20 paper/submitted/round2_png/intro_v4_cmame.tex
more paper changes
stashing edits

The choice of model should be informed by one's goal: for a range of applications, the parameters may be numerous, such as when corresponding to a discretized field, and yet what is ultimately of interest may be some low-dimensional quantity of interest (QoI). We consider this \textit{goal-oriented inverse problem}, where the goal of inferring parameters is to use them in predicting a QoI. In this context, one may choose a lower-fidelity model for the inference problem, sacrificing accuracy in the state and/or parameter estimates in exchange for reduced computational costs while keeping the error in the QoI to within some acceptable tolerance.  This paper formally poses the problem of managing the fidelity of modeling choices in solving the goal-oriented inverse problem, and develops an adjoint-based model adaptivity approach that achieves a desired level of accuracy in the QoI prediction. In particular, we adaptively form a mixed-fidelity model by using models of different levels of fidelity in different parts of the domain.

@@

Goal-oriented approaches prioritize accuracy in the QoI over accuracy in the states and/or parameters; one aims to reduce the cost of solving the problem (whether forward or inverse) while maintaining an acceptably low error tolerance for the QoI. In the context of the forward problem, goal-oriented methods have been developed for mesh refinement \cite{BecRann01,Park04,PrudOden99,VendDarm00,Yano12} and for model adaptivity~\cite{BraackErn03,OdenPrudetal06}. These methods derive an adjoint-based a posteriori error estimate for an output functional, and this estimate is used to guide adaptive mesh and/or model refinement.

@@

>>>>

The choice of model should be informed by one's goal: for a range of applications, the parameters may be numerous, such as when corresponding to a discretized field, and yet what is ultimately of interest may be some low-dimensional quantity of interest (QoI). We consider this \textit{goal-oriented inverse problem}, where the goal of inferring parameters is to use them in predicting a QoI. In this context, one may choose a lower-fidelity model for the inference problem, sacrificing accuracy in the state and/or parameter estimates in exchange for reduced computational costs while keeping the error in the QoI to within some acceptable tolerance. Compared to the highest-fidelity model, the lower-fidelity model could, for example, be a simplified model including fewer physical phenomena, be a reduced-order model, or have a reduced parameter space; this generalizes the setup where one only refines the mesh to create high-fidelity models. This paper formally poses the problem of managing the fidelity of modeling choices in solving the goal-oriented inverse problem, and develops an adjoint-based model adaptivity approach that achieves a desired level of accuracy in the QoI prediction. In particular, we adaptively form a mixed-fidelity model by using models of different levels of fidelity in different parts of the domain. %Compared to the high-fidelity model, inclusion of low-fidelity regions can make this mixed-fidelity model less nonlinear, and/or have a reduced state and/or parameter space;.

@@

Goal-oriented approaches prioritize accuracy in the QoI over accuracy in the states and/or parameters; one aims to reduce the cost of solving the problem (whether forward or inverse) while maintaining an acceptably low error tolerance for the QoI. In the context of the forward problem, goal-oriented methods have been developed for mesh refinement \cite{BecRann01,Park04,PrudOden99,VendDarm00,Yano12} and for model adaptivity~\cite{BraackErn03,OdenPrudetal06}. These methods derive an adjoint-based a posteriori error estimate for an output functional, and this estimate is used to guide adaptive mesh and/or model refinement. %CITE REVIEWER 4 PAPERS HERE (with braack and oden), SAY THAT ONE COULD USE GOAL-ORIENTED-FWD APPROACHES FOR INVERSE PROBLEMS BY RUNNING WITH FWD THAT IS KKT SYSTEM, BUT SAY THAT HERE TAKE ADVANTAGE OF THE PARTICULAR STRUCTURE ON AN INVERSE PROBLEM?

@@

\red{STILL NEED TO ADD REFERENCES TO THE TWO POINTED OUT BY REVIEWER 4}

### 2017-06-20 paper/submitted/round2_png/mathematical_formulation_v4_cmame_vikram_HLedit.tex
stashing edits

We assume that there exists an `ideal' model for solving the inverse problem, which we will call the high-fidelity model; given adequate computational resources, we would use this high-fidelity to formulate and solve our inverse problem. We use the subscript $_{HF}$ to denote the high-fidelity inverse problem,
%
\begin{subequations}
\label{eq:invOptHF}
\begin{align}
q_{HF},u_{HF} &=&\arg\min\limits_{q,u} & \quad J(q,u)=\frac{1}{2}\|d-C(u)\|_2^2 + R(q), \label{eq:invOpt_objHF} \\
&&\textrm{s.t. }& \quad a_{HF}(u,q)(\phi)=\ell_{HF}(q)(\phi),\quad\forall\phi\in U_{HF} \label{eq:invOpt_consHF}
\end{align}
\end{subequations}
%
where $q_{HF} \in Q_{HF}$ and $u_{HF}\in U_{HF}$ are the inferred parameters and corresponding state. In practice, one has limited computational resources, so one may need to instead infer the parameters using a less computationally demanding model, which we call the low-fidelity model. We use the subscript $_{LF}$ to denote the low-fidelity inverse problem,
%
\begin{subequations}
\label{eq:invOptLF}
\begin{align}
q_{LF},u_{LF} &=& \arg\min\limits_{q,u} & \quad J(q,u)=\frac{1}{2}\|d-C(u)\|_2^2 + R(q), \label{eq:invOpt_objLF} \\
&& \textrm{s.t. }& \quad a_{LF}(u,q)(\phi)=\ell_{LF}(q)(\phi),\quad\forall\phi\in U_{LF}, \label{eq:invOpt_consLF}
\end{align}
\end{subequations}
%
where $q_{LF} \in Q_{LF}$ and $u_{LF}\in U_{LF}$ are the inferred parameters and corresponding state. If there is \red{compatibility (need to elaborate on "compatibility" somehow; was this discussed in the stokes-NS paper? compatibility comes up in fusing the models, and having the same QoI functional apply to variables of all fidelities)} between the low-fidelity and high-fidelity solution spaces for the parameter and state variables, one may combine the two models to form a mixed-fidelity model and corresponding inverse problem. The low-fidelity model can be used in part of the computational domain, and the high fidelity in the remainder; we denote the mixed-fidelity model and its variables with the subscript $_{MF}$. How one chooses to combine the low- and high-fidelity models can be informed by what aspect of the inferred parameters is of interest.

In a {\em goal-oriented inverse problem}, the ultimate purpose of inferring the unknown parameters is to calculate some quantity of interest (QoI). We denote a scalar QoI by a functional that maps the parameters and state to our QoI, $I:Q \times U \to \R$. %The QoIs evaluated with the high ($I(q_{HF},u_{HF})$), low ($I(q_{LF},u_{LF})$) or mixed ($I(q_{MF},u_{MF})$) fidelity models correspondingly map parameters and state from the high, low or mixed fidelity function spaces to QoI values. 
We assume that the parameter and state spaces of the high-, low-, and mixed-fidelity models are such that the same QoI functional can be applied to each to obtain the corresponding QoI values. \red{Don't we also have to assume that the same regularization $R$ and observation $C$ can be applied? does that limit compatibility more than QoI?}
We consider the tradeoff between the error in the QoI and the fidelity (and corresponding computational expense) of the model we use, seeking computable estimates for the error incurred in approximating the high-fidelity model with lower fidelity models. Such an estimate allows us to adaptively choose regions where the high-fidelity model is used, starting from an initial low- or mixed- fidelity model.

>>>>

We assume that there exists an `ideal' model for solving the inverse problem, which we will call the high-fidelity model; given adequate computational resources, we would use this high-fidelity to formulate and solve our inverse problem. We use the subscript $_{HF}$ to denote the high-fidelity inverse problem,
%
\begin{subequations}
\label{eq:invOptHF}
\begin{align}
q_{HF},u_{HF} &=&\arg\min\limits_{q,u} & \quad J(q,u)=\frac{1}{2}\|d-C(u)\|_2^2 + R(q), \label{eq:invOpt_objHF} \\
&&\textrm{s.t. }& \quad a_{HF}(u,q)(\phi)=\ell_{HF}(q)(\phi),\quad\forall\phi\in U_{HF} \label{eq:invOpt_consHF}
\end{align}
\end{subequations}
%
where $q_{HF} \in Q_{HF}$ and $u_{HF}\in U_{HF}$ are the inferred parameters and corresponding state. In practice, one has limited computational resources, so one may need to instead infer the parameters using a less computationally demanding model, which we call the low-fidelity model. We use the subscript $_{LF}$ to denote the low-fidelity inverse problem,
%
\begin{subequations}
\label{eq:invOptLF}
\begin{align}
q_{LF},u_{LF} &=& \arg\min\limits_{q,u} & \quad J(q,u)=\frac{1}{2}\|d-C(u)\|_2^2 + R(q), \label{eq:invOpt_objLF} \\
&& \textrm{s.t. }& \quad a_{LF}(u,q)(\phi)=\ell_{LF}(q)(\phi),\quad\forall\phi\in U_{LF}, \label{eq:invOpt_consLF}
\end{align}
\end{subequations}
%
where $q_{LF} \in Q_{LF}$ and $u_{LF}\in U_{LF}$ are the inferred parameters and corresponding state. If there is \blue{compatibility (need to elaborate on "compatibility" somehow; was this discussed in the stokes-NS paper? compatibility comes up in fusing the models, and having the same QoI functional apply to variables of all fidelities) (just refer to section 3.1 here for more details on 'compatibility')} between the low-fidelity and high-fidelity solution spaces for the parameter and state variables, one may combine the two models to form a mixed-fidelity model and corresponding inverse problem. The low-fidelity model can be used in part of the computational domain, and the high fidelity in the remainder; we denote the mixed-fidelity model and its variables with the subscript $_{MF}$. How one chooses to combine the low- and high-fidelity models can be informed by what aspect of the inferred parameters is of interest.

In a {\em goal-oriented inverse problem}, the ultimate purpose of inferring the unknown parameters is to calculate some quantity of interest (QoI). We denote a scalar QoI by a functional that maps the parameters and state to our QoI, $I:Q \times U \to \R$. %The QoIs evaluated with the high ($I(q_{HF},u_{HF})$), low ($I(q_{LF},u_{LF})$) or mixed ($I(q_{MF},u_{MF})$) fidelity models correspondingly map parameters and state from the high, low or mixed fidelity function spaces to QoI values. 
We assume that the parameter and state spaces of the high-, low-, and mixed-fidelity models are such that the same QoI functional can be applied to each to obtain the corresponding QoI values. \blue{Don't we also have to assume that the same regularization $R$ and observation $C$ can be applied? does that limit compatibility more than QoI?}
We consider the tradeoff between the error in the QoI and the fidelity (and corresponding computational expense) of the model we use, seeking computable estimates for the error incurred in approximating the high-fidelity model with lower fidelity models. Such an estimate allows us to adaptively choose regions where the high-fidelity model is used, starting from an initial low- or mixed- fidelity model.
